---
title: "Project_1_607"
author: "Salma Elshahawy"
date: "9/16/2019"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Chess is known as a noble game – “a game of kings“, and it is no coincidence, as it takes/develops similar skills. 

In chess, just like in life, you can either aimlessly wonder around, hoping for a miracle, or set yourself a goal, build a plan and strive to accomplish it. The latter is no guarantee of success, but the former is a sure recipe for failure. **Chess teaches you to have a goal and overcome the obstacles on your way there**. 

## Requirements

In this project, a text file with chess tournament results where the information has some structure given to me - in a text format. 
The requirements were to :

* Create an R Markdown file that generates a .CSV file that could for example be imported into a SQL database. 
* The information should be like the following :
  
  Player's Name | Player's State | Total Number of Points | Player's Pre-Rating | Average Pre Chess Rating of Opponnents
  --------------|----------------|------------------------|---------------------|----------------------------------------
  Gary Hua      | ON             | 6.0                    | 1794                |  1605
  
## Questions 

For this project, I was curious to get more insights into thses dataset. 
1. Who has the most Pre-rating? 
2. Which state has players with highest pre-rating? 
3. What are the factors affects on winning tournaments? 

## Work wireframe 

I decided to put a plan how do I will handle, analyze the data given. This plan breaks down into steps, so I can easily follow these steps for more organized work. 

## Steps to follow (Methodology): 

1. Save the original tournamentinfo.txt file into my local machine then push it to my personal GitHub account. 
2. Download necessary libraries.
3. Read the file as a dataframe using R Markdown.  
4. Clean up the data from any special characters, empty strings,..etc.
5. Combine each two consecutive rows into one row - The table in the txt file are with **two tier headers and rows** 
6. Seperate the columns from one column into multiples. 
7. Get the opponents pre-rating year for each player 
8. Calculate the average pre-rating for all opponents pre-rating for each player.
9. Explore the final dataframe and write the table to a csv file.
10. Refactor repeatitive code into functions, Make a **DRY code**
10. Do analysis to the defined model and get conclusion.


### Loading necessary liberaries 

```{r}
library(RCurl)
library(htmlTable)
library(magrittr)
library(kableExtra)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(Hmisc)
```

### Here I pushed the file into my GitHub account and store it into a variable and call read.delim() on it.

```{r}
url <- 'https://raw.githubusercontent.com/salma71/MSDS_2019/master/Fall2019/aquisition%26management_607/project_1/tournamentinfo.txt'

# reading the url as a dataframe
tourn <- read.delim(url, header = FALSE, stringsAsFactors = FALSE)
head(tourn) %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```
```{r}
is.data.frame(tourn)
```

As shown in the output, the table has many issues need to be handle first before analysing: 
  * The table has (-) needs to be removed
  * Has an empty rows
  * The data of each player is devided into two rows. 

```{r}
# remove the ---- using str_remove_all with a regex pattern

tourn_2 <- data.frame(str_remove_all(tourn$V1,"(\\-+)"))
head(tourn_2) %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```


```{r}
# now removing the extra empty rows from all the dataframe

tourn_2 <- data.frame(tourn_2[!apply(tourn_2 == "", 1, all),])
tourn_2 %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```

Now to the hard part, the target is to loop over the **tourn_2** and merge every two consecutive rows into one single row. To acheive that approach, I initialized an empty dataframe and store it into a variable called **new_table**. After that, let the for-loop iterate over the dataframe dimension of **6 X 1**, if the row number is odd, then seperate into two parts. Finally push them into the **New_table** dataframe. 

```{r}
# Need to define an empty new_dataframe
new_table <- data.frame(c())
# Combining two consecutive rows into one column
for (i in 1:dim(tourn_2)[1]){
  if (i %% 2 != 0) {
    Part1 <- rbind(new_table$Part1, as.character(tourn_2[i,1]))
    Part2 <- as.character(tourn_2[i+1,1])
    binded <- data.frame(paste0(Part1, Part2))
    names(binded) <- "Binded"
    new_table <- rbind(new_table, binded)
  } 
}
head(new_table) %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="120%",height="300px")
```

```{r}
# check that the new generated new_table is a dataframe
is.data.frame(new_table)
```


Next step, we need to chop this long column into multiple columns, each column represents piece of data. For this purpose, I used built in method within tidyverse library called separate() which takes three arguments which are dataframe, the column that needs to be chopped, new columns names, and finally the separator where the split should be done. For this case, I wanted t split on the "|". 

For the columns names, I put it into a vector and stored it in a variable called "into"
```{r}
into <- c("Pair", "Player Name","Total","Round","Round.1","Round.2","Round.3","Round.4","Round.5","Round.6","State","USCF ID / Rtg (Pre>Post)","Pts","1","2", "3","4","5","6", "7")
splitted_df <- separate(new_table, Binded, into, sep = "\\|")
splitted_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")

```

```{r}
colnames(splitted_df)
```
Now the data has been successfully splitted into **21** columns. However, the first row is a duplicated header. In this case, this row can be deleted. 
```{r}
clean_df <- splitted_df[-c(1),]
clean_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")
```

```{r}
colnames(clean_df)
```

Now, we need to split the **clean_df** again at column **"USCF ID / Rtg (Pre>Post)"** to get pre and post rating columns to get the average as required at the beginning of the aasignment. To do so, I used Regex patterns to remove punctuations and generate three different columns for USCF ID, Pre, and Post rating using tidyverse method **separate()**. 


```{r}
rtg_extract <- str_extract_all(clean_df$`USCF ID / Rtg (Pre>Post)`, "\\b\\d{1,}")
#as.data.frame(rtg_split)
colnames(rtg_extract, do.NULL = FALSE) 

rtg_df <- data.frame(as.character(str_remove_all(rtg_extract, "[[:punct:],c]")))

header_name <- c("USCF ID", "Pre", "Post")
# rename columns
colnames(rtg_df)
names(rtg_df)[names(rtg_df) == "as.character.str_remove_all.rtg_extract......punct...c...."] <- "years"

rtg_df <- separate(rtg_df, years, header_name, sep = " ") 
rtg_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```

Now we need to merge those three columns into the main dataframe **clean_df**. Before doing the merge, we need to drop the "USCF ID / Rtg (Pre>Post)" column from **clean_df** then merge. 

```{r}
clean_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")
```

```{r}
# get the column number => 12
colnames(clean_df)
```

```{r}
# delete column 12
drop_uncleandata <- clean_df[,-c(12)]
drop_uncleandata %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")

```

Merge the new three columns into **drop_uncleandata** 

```{r}
tournament_df <- cbind(drop_uncleandata, rtg_df)
tournament_df  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")
```
 
```{r}
# get columns names and positions
colnames(tournament_df)
```

```{r}
#re-order the columns name 
ordered_tournament_df <- tournament_df[c(1,2,11,3,21,22,12,4,5,6,7,8,9,10,13,14,14,16,17,18,19,20)]
ordered_tournament_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="150%",height="300px")
```

It is better to take a subset from the data to focus on it and make plausible analysis into it. So, I took a subset from column 1 to column 7. Then, I add a new column for average initialized with 0 value - this will be the calculated average pre-rating for each player. 

```{r}
final_table <- subset(ordered_tournament_df, select = c(1:7))
head(final_table)  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```
Initialize **avg_rtg** varaible filled with 0 numeric value.

```{r}
final_table$avg_rtg <- 0
final_table %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```


```{r}
#score table - a subset only has pair and player's name
score_df <- subset(ordered_tournament_df, select = c(1:2))
score_df

```

Now, Creating the opponents dataframe, to get number of games, and pre- rating for each player and opponent.

```{r}
opponent1 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round`,"[[:digit:]]{1,}")))
opponent2 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.1`,"[[:digit:]]{1,}")))
opponent3 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.2`,"[[:digit:]]{1,}")))
opponent4 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.3`,"[[:digit:]]{1,}")))
opponent5 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.4`,"[[:digit:]]{1,}")))
opponent6 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.5`,"[[:digit:]]{1,}")))
opponent7 <- data.frame(as.numeric(str_extract_all(ordered_tournament_df$`Round.6`,"[[:digit:]]{1,}")))

# Creating Opponents data frame.
opponents <- cbind(score_df, opponent1, opponent2, opponent3, opponent4, opponent5, opponent6, opponent7)

# Finding number of games played
for(i in 1:dim(opponents)[1]){
    opponents$NGames[i] <- 7 - as.numeric(sum(is.na(opponents[i,])))
}

# Reporting  to view the opponents table

names(opponents) <- c("Pair","Player Name","Opp 1","Opp 2","Opp 3","Opp 4","Opp 5","Opp 6","Opp 7", "NGames")
opponents  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```
More cleaning up - NA -> 0 value.

```{r}
# Eliminating NA Cases in order to continue with our calculations, NA replaced by 0.
opponents[is.na(opponents)] <- as.numeric(0)

opponents  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```

Merging **final_df** with the opponents 
```{r}
final_df <- merge(final_table, opponents)

final_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")


```
Calculating the average pre-rating for each player. Construct a for-loop to iterate over the dataframe and if the value of the opponent is equal to any pair number in the df then calculate the average value. 

```{r}
# Procedure to calculate Average Pre-Rating for each player

for (k in 3:9){
  for (j in 1:dim(final_table)[1]){ 
    for (i in 1:dim(final_table)[1]){
      if (as.numeric(opponents[j,k]) == as.numeric(final_table$Pair[i])){
        final_table$avg_rtg[j] <- as.numeric(final_table$avg_rtg[j]) + as.numeric(final_table$Pre[i]) 
      }
    }
  }
}


final_table$avg_rtg <- round(as.numeric(final_table$avg_rtg) / as.numeric(opponents$NGames),0)

final_table  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```
## Making analysis

Need to know the which player has the max average pre-rating compared to his peers aggregated by state. Bar chart are the best in describing that. 

To analyse, I will merge **final_table** with **final_df** in one table to get insight details.

```{r}
chess_df <- merge(final_table, final_df, by=c("Pair", "Player Name", "State", "Pre", "Post", "Total", "Pts"))
colnames(chess_df)
chess_df[,-c(9:16)]
chess_df
```
```{r}
summary(chess_df$avg_rtg.x)
```
How many players per state involved at this tournmant. Count observations by group is always a good idea. We can aggregate the the number of occurence with n(). The code below computes the number of Players played by State.

```{r}
# count observations
data_df <- chess_df %>%
	group_by(State) %>%
	summarise(`Pair` = n())
data_df
ggplot(data = data_df,aes(x = State, y = `Pair`)) +
    geom_col(aes(fill = State), position = "dodge") + 
    geom_text(
    aes(label = `Pair`, y = `Pair` + 0.7),
    position = position_dodge(0.9),
    vjust = 0
  )

```

As concluded from the graph that the majority of the players are from **MI** where 85% of players are from there. 

Now, let's check out if **MI** residents plays all rounds. 

```{r}
chess_df %>% 
filter(NGames >= 1) %>%
ggplot(aes(x = State, y = avg_rtg.x, colour = NGames)) +
geom_point(alpha = 0.6,  position = position_jitter()) + stat_smooth(method = "lm")
```
It shows that the majority of players in the three states played their 7 games. 

Group by **average pre-rating** for each player per **State** to find out each state statistics (mean, median, max, and min)

```{r}
ggplot(chess_df, aes(x = State, y = avg_rtg.x)) +
  geom_boxplot() +
    stat_summary(
      fun.y = mean, 
      fun.ymin = min, 
      fun.ymax = max,
      colour = "blue",
      size = 1
    )
```
As Illustrated, the mean for **ON** state is higher than **MI** state, the max is higher fir **ON** as well; however, the min is nearly the same. From observation, we can exclude **OH** from comparasion because it has only one player and the pre-rating average value is above **1500**.

```{r}
chess_df %>%
filter(avg_rtg.x >= 1400) %>%
ggplot(aes(`Player Name`, avg_rtg.x)) +
  geom_label(aes(label = `Player Name`), size = 2) + 
  facet_grid( ~ State )
  
```
As shown in the label graph that although **ON** has less players than **MI**, it has the highest pre-rating average compared to the two other states. **GARY HUA** has the highest average of **1600**.


Group by **Player Name** and **State** to find out each state statistics (mean, median, max, and min)

_________________________________________

## Adjusting the table header to exactly matches the user requirement.

In this section, I will remove the extra columns for Pair,Post, Pts and change the columns names to match the user requirements .

```{r}
colnames(chess_df)
```

```{r}
chess_tournament <- chess_df[, -c(1,5,7,9,10,11,12,13,14,15,16,17)]
chess_tournament
names(chess_tournament) <- c("Player’s Name", "Player’s State", "Total Number of Points", "Player’s Pre-Rating", " Average Pre Chess Rating")
chess_tournament
```
## Export to CSV file 

```{r}
file_to_export <- write.table(chess_tournament, file = "chess_tournamentInfo.csv", row.names = FALSE, na="", col.names = TRUE, sep = ",")

```
## Test that the file is already created. 

*Please be sure that you set working directory in Rstudio to the current working directory.*

```{r}
file_test("-f", "~/Desktop/MSDS_2019/Fall2019/aquisition&management_607/project_1/chess_tournamentInfo.csv")
```

## open the file from local machine

```{r}
file_to_open <- read.csv(file = "chess_tournamentInfo.csv", header = TRUE, sep = ",")

file_to_open   %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```

## Open the file from GitHub account 

```{r}
url <- 'https://raw.githubusercontent.com/salma71/MSDS_2019/master/Fall2019/aquisition%26management_607/project_1/chess_tournamentinfo.csv'

# reading the url as a dataframe
tourn <- read.delim(url, header = FALSE, stringsAsFactors = FALSE)
head(tourn) %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```


## Conclusion and next steps

After getting basic exploring analysis for the chess tournament, we need further information about each player to get more details about the winning factors. For example, have more details about game time for each player, age, education,..etc. 


