---
title: "607_week_2_assignment"
author: "Salma Elshahawy"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## SQL and R 

In this report, a basic survey was performed to rate the most recent top Box Office movies. The sample taken is six-6 movies from different Genres. The basic survey taken from friends and family members. 

#### Technologies used: 
  1. PostgeSQL Database.
  2. PgAdmin 4 software to create/execute sql scripts.
  3. RPostgreSQL pakage for database connection.
  4. KableExtra for styling the output tables.

#### The table was created and filled with data using PgAdmin software, the SQL script is attached to this assignment on the blackboard portal. 
______________________________________________________________

### First step: Connect DB - PostgreSQL to RStudio.

```{r}
#Connect postgresql db to Rstudio
require("RPostgreSQL")

pw <- {
  "yahia2010"
}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "movies_rating",
                 host = "localhost",
                 port = 5432,
                 user = "postgres",
                 pw)
# clearing the pw 
rm(pw)
#check if the database is connected
dbExistsTable(con, "movie_rating")

```


```{r}
library(kableExtra)

# return the table 
dtab = dbGetQuery(con, "select * from movie_rating")

dtab %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```

```{r}
## here we have to clean the data set first before we start to analyise it.
str(dtab)
```

```{r}
#The results of complete.cases() is a logical vector with the value TRUE for rows that are complete, and FALSE for rows that have some NA values. To remove the rows with missing data
complete.cases(dtab)
```


```{r}
# Here we delete the last two rows from the dataset
#clean_data <- dtab[complete.cases(dtab), ]
#str(clean_data)

## we can use na.omit() function to omit all rows that contain NA values
clean_data <- na.omit(dtab)
clean_data  %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```

```{r}
## Building a dotplot graph to illustrate the data
require(ggplot2)
ggplot(clean_data, aes(x = movie_rating, y = movie_name)) + geom_boxplot() + theme_bw()
```

```{r}
# close the connection
dbDisconnect(con)

```

