---
title: "607_week_2_assignment"
author: "Salma Elshahawy"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## SQL and R 

In this report, a basic survey was performed to rate the most recent top Box Office movies. The sample taken is six-6 movies from different Genres. The basic survey taken from friends and family members. 

#### Technologies used: 
  1. PostgeSQL Database.
  2. PgAdmin 4 software to create/execute sql scripts.
  3. RPostgreSQL pakage for database connection.
  4. KableExtra for styling the output tables.

### First step: Connect DB - PostgreSQL to RStudio.

```{r}
#Connect postgresql db to Rstudio
require("RPostgreSQL")

pw <- {
  "yahia2010"
}
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "movies_rating",
                 host = "localhost",
                 port = 5432,
                 user = "postgres",
                 pw)
# clearing the pw 
rm(pw)
dbExistsTable(con, "movie_rating")

```


```{r}
library(kableExtra)

movie_rating_df = read.csv("/Users/salmaelshahawy/Desktop/MSDS_2019/Fall2019/aquisition&management_607/week_2/movie_rating", header = TRUE) 

movie_rating_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```

```{r}
dim(movie_rating_df)
```

