---
title: "Elshahawy-project"
author: "Salma Elshahawy"
date: "10/12/2019"
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
  html_document:
    code_folding: show
    df_print: paged
    highlight: pygments
    mathjax: local
    number_sections: no
    pdf_document: default
    rmarkdown::html_document: null
    theme: cosmo
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, warning=FALSE, message=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part 1 - Introduction

I decided to use the dataset of nycflights13::flights: package included with R. This package contains information about all flights that departed from NYC (i.e., EWR, JFK and LGA) in 2013: 336,776 flights with 16 variables. To help understand what causes delays, it also includes a number of other useful datasets: weather, planes, airports, airlines. [Source: Bureau of transportation statistics](http://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=236)

H0(null hypotithes) -> No associations between departure delay and arrival delay

HA(alternative hypotithes) -> There are associations between departure delay and arrival delay.

+ Research Questions:

  + Are the actual departure delay associated with the arrival delay? 



### Part 2 - Data

+ Variables:

  + variable_1 -> dep_delay - independent variable, numerical - categorical

  + outcome -> arr_delay, numerical - numerical - categorical

+ There are about 336,776 observation in the given dataset. Each observation represent flight full details.

+ This is an observational study. I will draw my conclusions based on analyzing the existing data.


### Part 3 - Exploratory data analysis

```{r}
library(RCurl)
library(dplyr)
library(ggplot2)
library(ggExtra)

library(nycflights13)
head(flights) 
```


```{r}
summary(flights)
```



```{r}

# taking a subset
sub_set <- flights[c(6,9,10,16)]
sub_set

summary(sub_set)
```

```{r}
## get statistical analysis for the whole population 
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(sub_set, aes(arr_delay, dep_delay)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
```


```{r}

## sampling, get summary plots --> statistics for only sample of 100

sample_100 <- sample_n(sub_set, 100)

summary(sample_100)

theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(sample_100, aes(arr_delay, dep_delay)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
```

Both dep_delay and arr_delap are right skewed distribution. 


```{r}

theme_set(theme_bw())  # pre-set the bw theme.
ggplot(sample_100, aes(dep_delay, arr_delay)) + 
  geom_jitter(aes(colour = dep_delay, size = arr_delay), na.rm = TRUE) +
  coord_flip()
  
```



### Part 4 - Inference


This dataset doesn't follow the normal distribution. Since n = 100 => which is more than 25 we can do a linear regression model. Let's begin with the correlation which is a statistical tool to measure the level of linear dependence between two variables, that occur in pair  

```{r}
cor(sub_set$arr_delay, sub_set$dep_delay, use = "complete.obs")
```

The correlation is very strong as it close to 1 - strong correlation. Now, let's build the linear regression model.

```{r}
linearMod <- lm(arr_delay ~ dep_delay, data=sub_set)  # build linear regression model on full data
print(linearMod)
```

Now that we have built the linear model, we also have established the relationship between the predictor and response in the form of a mathematical formula for arrival delay (arr_delay) as a function for distance. For the above output, you can notice the ‘Coefficients’ part having two components: Intercept: -5.899, distance: 1.019 These are also called the beta coefficients. In other words,

arr_delay = Intercept + (beta ∗ dep_delay)

```{r}
summary(linearMod)  # model summary
```

  arr_delay = -5.899 + 1*dep_delay

### Part 5 - Conclusion

as a conclusion, I would go with refusing the Null hypotethis that there is no associations between arrival delay and departure delay. 

### References

[Flights database](http://blog.rstudio.org/2014/07/23/new-data-packages/)
