---
title: "Applying linear regression to study flights delay"
author: "Salma Elshahawy"
date: "10/12/2019"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: pygments
    number_sections: yes
    rmarkdown::html_document: null
    theme: cosmo
    toc: yes
    toc_float: yes
  pdf_document:
    latex_engine: xelatex
    toc: yes
---

```{r setup, include=FALSE, warning=FALSE, message=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1 - Introduction

I decided to use the dataset of nycflights13::flights: package included with R. This package contains information about all flights that departed from NYC (i.e., EWR, JFK and LGA) in 2013: 336,776 flights with 16 variables. To help understand what causes delays, it also includes a number of other useful datasets: weather, planes, airports, airlines. [Source: Bureau of transportation statistics](http://www.transtats.bts.gov/DL_SelectFields.asp?Table_ID=236)

H0(null hypotithes) -> No associations between departure delay and arrival delay

HA(alternative hypotithes) -> There are associations between departure delay and arrival delay.

+ Research Questions:

  + Are the actual departure delay associated with the arrival delay? 



## Part 2 - Data

+ Variables:

  + variable_1 -> dep_delay - independent variable, numerical - discrete

  + outcome -> arr_delay, numerical - numerical - discrete

+ There are about 336,776 observation in the given dataset. Each observation represent flight full details.

+ This is an observational study. I will draw my conclusions based on analyzing the existing data.


## Part 3 - Exploratory data analysis

```{r, message = FALSE, warning=FALSE}
library(RCurl)
library(dplyr)
library(ggplot2)
library(ggExtra)

library(nycflights13)
flights
```


```{r, message = FALSE, warning=FALSE}
summary(flights)
```



```{r, message = FALSE, warning=FALSE}

# taking a subset
sub_set <- flights[c(6,9,10,16)]
sub_set

summary(sub_set)
```

```{r, message = FALSE, warning=FALSE}
## get statistical analysis for the whole population 
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(sub_set, aes(dep_delay, arr_delay)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
```


```{r, message = FALSE, warning=FALSE}

## sampling, get summary plots --> statistics for only sample of 100

sample_100 <- sample_n(sub_set, 100)

summary(sample_100)

theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(sample_100, aes(dep_delay, arr_delay)) + 
  geom_count() + 
  geom_smooth(method="lm", se=F)

ggMarginal(g, type = "histogram", fill="transparent")
```

Both dep_delay and arr_delap are right skewed distribution. 


```{r, message = FALSE, warning=FALSE}

theme_set(theme_bw())  # pre-set the bw theme.
ggplot(sample_100, aes(dep_delay, arr_delay)) + 
  geom_jitter(aes(colour = dep_delay, size = arr_delay), na.rm = TRUE) 
  
```



## Part 4 - Inference


This dataset doesn't follow the normal distribution. Since n = 100 => which is more than 25 we can do a linear regression model. Let's begin with the correlation which is a statistical tool to measure the level of linear dependence between two variables, that occur in pair  

```{r message=FALSE, warning=FALSE}
library(GGally)
sample_100 <- sample_100 %>%
  na.omit() %>%
  select(dep_delay, arr_delay, distance)

ggpairs(data = sample_100, title = "title")
# cor(sub_set$arr_delay, sub_set$dep_delay, use = "complete.obs")
```

The correlation between arr_delay and dep_delay is very strong as it close to 1 - strong correlation. However, relation doesn't mean causation. Now, let's build the linear regression model.

```{r, message = FALSE, warning=FALSE}
linearMod <- lm(arr_delay ~ dep_delay, data=sample_100)  # build linear regression model on full data
summary(linearMod)
```

### Check residuals distribution

This plot shows if residuals are normally distributed. Do residuals follow a straight line well or do they deviate severely? It’s good if residuals are lined well on the straight dashed line. 


```{r, message = FALSE, warning=FALSE}
library(DATA606)
plot_ss(sample_100$dep_delay, sample_100$arr_delay)
```



```{r, message = FALSE, warning=FALSE}
fit <- linearMod
d <- sample_100
d$predicted <- predict(fit)   # Save the predicted values
d$residuals <- residuals(fit) # Save the residual values
ggplot(d, aes(x = dep_delay, y = arr_delay)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +     # regression line  
  geom_segment(aes(xend = dep_delay, yend = predicted), alpha = .2) +      # draw line from point to line
  geom_point(aes(color = abs(residuals), size = abs(residuals))) +  # size of the points
  scale_color_continuous(low = "green", high = "red") +             # colour of the points mapped to residual size - green smaller, red larger
  guides(color = FALSE, size = FALSE) +                             # Size legend removed
  geom_point(aes(y = predicted), shape = 1) +
  theme_bw()
```

```{r }
plot(fit, which=1, col=c("blue"))
```

When we look at the plot above, we see that the data does not have any obvious distinct pattern. While it is slightly curved, it has equally spread residuals around the horizontal line without a distinct pattern.

This is a good indication it is not a non-linear relationship.


```{r, message = FALSE, warning=FALSE}
plot(fit, which=2, col=c("blue")) # qqplot
```

For our model, the Q-Q plot shows pretty good alignment to the the line with a few points at the top slightly offset. Probably not significant and a reasonable alignment.

```{r, message = FALSE, warning=FALSE}
plot(fit, which=3, col=c("blue"))  # Scale-Location Plot
```

The residuals are reasonably well spread above and below a pretty horizontal line however the beginning of the line does have more points so has less variance there.

Residual analysis plots are a very useful tool for assessing aspects of veracity of a linear regression model on a particular dataset and testing that the attributes of a dataset meet the requirements for linear regression. 

******************************

Now that we have built the linear model, we also have established the relationship between the predictor and response in the form of a mathematical formula for arrival delay (arr_delay) as a function for departure delay. For the above output, we can notice the ‘Coefficients’ part having two components: Intercept: -6.94, distance: 1.019 These are also called the beta coefficients. In other words,

arr_delay = Intercept + (beta ∗ dep_delay)

arr_delay = -5.899 + 1.02*dep_delay


## Part 5 - Conclusion

as a conclusion, I would go with refusing the Null hypotethis that there is no associations between arrival delay and departure delay. However, We need to consider other attributes that has a confounding effects on the arrival times.

## References

+ [Flights database](http://blog.rstudio.org/2014/07/23/new-data-packages/)

+ [University of Virginia Library](http://data.library.virginia.edu/diagnostic-plots/)
