---
title: "week_3_assignment_607"
author: "Salma Elshahawy"
date: "9/10/2019"
output:
  html_document:
    rmarkdown::html_document:
    code_folding: show
    df_print: paged
    highlight: pygments
    number_sections: no
    pdf_document: default
    theme: lumen
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Chapter_8 :Automated Data Collection in R

## Problem_3

Copy the introductory example. The vector name stores the extracted names.

```{r}
#R> name
     #[1] "Moe Szyslak"       "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
     #[4] "Ned Flanders"      "Simpson, Homer"       "Dr. Julius Hibbert"
```

##### a. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r}
# we create a variable to store the raw data given in the problem
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

raw.data
```

```{r}
# 1. we need to extract the alphabetic characters, we use [[:alpha:]., ]
# 2. we add a qualifier to impose the restriction that contents of the character class have to be matched at least twice to be considered a match. 
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names
```

```{r}
# check if names is df
is.data.frame(names)
```

```{r}
# convert names to dataframe vector
require(devtools)
library(kableExtra)

names_df <- as.data.frame(names)
names_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")

```

```{r}
# loop over the vector elements and if the the name contains ',' then make the first last and vise versa else keep the the string the same
splits <- str_split_fixed(names_df$names, ",", 2)

splits

```

```{r}
# rejoin the 2 columns to get proper names
names_df <- paste(splits[,2], splits[,1], sep = ' ')
names_df %>% kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width="100%",height="300px")
```

##### b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r}
# I used the built in function within stringr package str_detect witha a regex pattern to detect the title.

has_title <- str_detect(names_df, "^\\s[A-Za-z]{2,3}\\W+\\B\\s")
has_title
```

##### c. Construct a logical vector indicating whether a character has a second name.

```{r}
# I used the same built in function to see if there is a match for S. name pattern using regex
has_second_name <- str_detect(names_df, "[A-Z]{1}\\.\\s")
has_second_name
```

## Problem_4 

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

##### a. "[0-9]+\\$" 
- type: This is a numeric type \n
- from 0 to 9 match any charcter, case  sensitive \n
- the + sign is the quantifier means match one or more from the preceding token.\n
- the \$ the excape character match $ sign.\n
- example: 500$\n

##### b. "\\b[a-z]{1,4}\\b"
- type: This is a alphabetic type \n 
- "b" means that this is the word boundry \n
- "[a-z]" alphabetic characters from a to z \n
- "{1,4}" the qualifier match one to 4 times of the preceding token \n
- example: "bees" and "sara" \n

##### c. ".*?\\.txt$"

- type: Mixed with numeric, alphabetic, and special characters - more general pattern \n
- "." matches any char except the line breaks \n
- "*" the quantifier matches 0 or more from the preceding token. \n
- "?" lazy token makes the preceding quantifier lazy - matching as few chars as possible. \n
- "txt" match exactly 3 chars t, x, and t following the same order. \n
- "$" end of the string or end of the line. \n 
- example: " c.txt", "a.txt", "        a.txt" \n

##### d. "\\d{2}/\\d{2}/\\d{4}"

- type: This a numeric characters - most likely only integers of 7 digits \n
- "d" matches any digit \n
- "{2}" the quantifier matches 2 from the preceding token. \n
- "{4}" the quantifier matches 4 from the preceding token. \n
- example: "23984560" and "20708090" \n

##### e. "<(.+?)>.+?</\\1>"

- type: This a mixed type numeric alphabetic characters - most likely used for selecting DOM(HTML) tags
- "<" matches the char of < \n
- "()" means a group of regex. \n
- "." matches any char except line break. \n
- "+" is the quantifier means match one or more from the preceding token.\n
- "?" lazy token makes the preceding quantifier lazy - matching as few chars as possible. \n
- ">" matches the char of >
- example: an HTML tag "<a>hi<a>" and "< a>hi< a>" \n

## Problem_9

The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! 

> clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
     Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
     d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
     fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr
     
```{r}
secret_msg = "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
     Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
     d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
     fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
# First, I extracted all the upper case letters from the text and the dots
msg <- str_extract_all(secret_msg, "[[:upper:][:punct:]]{1,}")
# converting the string into a vector of chars
msg_df <- unlist(msg)
msg_df
# concatenate all the chars with each other
msg_df <- str_c(msg_df, collapse = "")
# replace all the dots with white spaces
msg <- str_replace_all(msg_df, "\\.", " ")
msg

```


